Monitoring Time Series With Missing Values:\\a Deep Probabilistic Approach

### 2022-02-17 paper.tex
added example

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{amsmath,amsfonts,graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary {shapes.geometric}
\usetikzlibrary {positioning}

@@

\begin{figure}
\begin{subfigure}{.64\textwidth}
\begin{tikzpicture}
\node [rectangle, draw] (rnn) {RNN};

@@

We used a 3-layer GRU-based recurrent neural network with hidden
size 64 and 20\% dropout between layers. We trained the network
with lookahead depths (number of steps to forecast in the
future) 2, 4, 8, and 16 using the Adam optimizer with learning
rate 0.001, training for 20 epochs (sufficient for convergence).
We performed the training on a cloud computing node with 1
NVIDIA T4 GPU, 4 Intel Xeon Platinum CPUs, and 64 Gb memory. The
training of a single model took 20 minutes.

We compared our approach with conventional imputation methods
`replace by the mean' and `replace by a random sample'.
In the `replace by mean' method, a missing value is replaced 
by the mean of the forecast. In the `replace by a random
sample' method, a missing value is replaced by a random sample
drawn from the forecast. As a performance metrics, we
used per-point negative log-likelihood loss on the test set. 
Tables~\ref{tab:replace-by-mean} and~\ref{tab:replace-by-sample}
show the difference in loss between uncertainty propagation and
`replace by the mean' and `replace by a random sample',
correspondingly. The greater is the number, the worse is the
forecasting by each of the methods compared to uncertainty
propagation.
\begin{table}[t]
\caption{Uncertainty propagation vs. `replace by the mean`.}
\label{tab:replace-by-mean}
	\setlength\tabcolsep{18pt}
	\centering
	\begin{tabular}{r|c c c c}
	 \textbf{missing} & \textbf{2} & \textbf{4} & \textbf{8} & \textbf{16} \\ \hline
		  \textbf{5\%} & 0.001 & 0.001 &  0.06 & 0.10  \\
		 \textbf{10\%} & 0.001 & 0.002 &  0.08 & 0.11  \\
		 \textbf{20\%} & 0.003 & 0.003 &  0.11 & 0.13  \\
		 \textbf{50\%} & 0.004 & 0.006 &  0.12 & 0.16  
	\end{tabular}
\end{table}
\begin{table}[t]
\caption{Uncertainty propagation vs. `replace by a random sample`.}
\vspace{1em}
\label{tab:replace-by-sample}
	\setlength\tabcolsep{18pt}
	\centering
	\begin{tabular}{r|c c c c}
	 \textbf{missing} & \textbf{2} & \textbf{4} & \textbf{8} & \textbf{16} \\ \hline
		  \textbf{5\%} & 0.04 & 0.04 & 0.13 & 0.14 \\
		 \textbf{10\%} & 0.06 & 0.06 & 0.16 & 0.24 \\
		 \textbf{20\%} & 0.11 & 0.12 & 0.20 & 0.27 \\
		 \textbf{50\%} & 0.18 & 0.19 & 0.28 & 0.30   
	\end{tabular}
\end{table}
One can see that in all cases uncertainty propagation provides
better forecasts than either of the conventional methods.

>>>>

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{amsmath,amsfonts,graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary {shapes.geometric}
\usetikzlibrary {positioning}

@@

\begin{figure}[t]
\begin{subfigure}{.64\textwidth}
\begin{tikzpicture}
\node [rectangle, draw] (rnn) {RNN};

@@

\begin{table}[t]
\caption{Uncertainty propagation vs. `replace by the mean`.}
\label{tab:replace-by-mean}
	\setlength\tabcolsep{18pt}
	\centering
	\begin{tabular}{r|c c c c}
	 \textbf{missing} & \textbf{2} & \textbf{4} & \textbf{8} & \textbf{16} \\ \hline
		  \textbf{5\%} & 0.001 & 0.001 &  0.06 & 0.10  \\
		 \textbf{10\%} & 0.001 & 0.002 &  0.08 & 0.11  \\
		 \textbf{20\%} & 0.003 & 0.003 &  0.11 & 0.13  \\
		 \textbf{50\%} & 0.004 & 0.006 &  0.12 & 0.16  
	\end{tabular}
\end{table}
\begin{table}[t]
\caption{Uncertainty propagation vs. `replace by a random sample`.}
\vspace{1em}
\label{tab:replace-by-sample}
	\setlength\tabcolsep{18pt}
	\centering
	\begin{tabular}{r|c c c c}
	 \textbf{missing} & \textbf{2} & \textbf{4} & \textbf{8} & \textbf{16} \\ \hline
		  \textbf{5\%} & 0.04 & 0.04 & 0.13 & 0.14 \\
		 \textbf{10\%} & 0.06 & 0.06 & 0.16 & 0.24 \\
		 \textbf{20\%} & 0.11 & 0.12 & 0.20 & 0.27 \\
		 \textbf{50\%} & 0.18 & 0.19 & 0.28 & 0.30   
	\end{tabular}
\end{table}
We used a 3-layer GRU-based recurrent neural network with hidden
size 64 and 20\% dropout between layers. We trained the network
with lookahead depths (number of steps to forecast in the
future) 2, 4, 8, and 16 using the Adam optimizer with learning
rate 0.001, training for 20 epochs (sufficient for convergence).
We performed the training on a cloud computing node with 1
NVIDIA T4 GPU, 4 Intel Xeon Platinum CPUs, and 64 Gb memory. The
training of a single model took 20 minutes.

We compared our approach with conventional imputation methods
`replace by the mean' and `replace by a random sample'.
In the `replace by mean' method, a missing value is replaced 
by the mean of the forecast. In the `replace by a random
sample' method, a missing value is replaced by a random sample
drawn from the forecast. As a performance metrics, we
used per-point negative log-likelihood loss on the test set. 
Tables~\ref{tab:replace-by-mean} and~\ref{tab:replace-by-sample}
show the difference in loss between uncertainty propagation and
`replace by the mean' and `replace by a random sample',
correspondingly. The greater is the number, the worse is the
forecasting by each of the methods compared to uncertainty
propagation.
One can see that in all cases uncertainty propagation provides
better forecasts than either of the conventional methods.

\begin{figure}[t]
\includegraphics[width=0.9\linewidth]{dist-vs-mean.pdf}
\caption{Uncertainty propagation vs `replace by mean'. 95\% confidence intervals are shaded.}
\label{fig:dist-vs-mean}
\end{figure}
As an illustration of the advantage of uncertainty propagation,
consider Figure~\ref{fig:dist-vs-mean}, which shows forecasts
using uncertainty propagation and `replace by mean' in presence
of missing values. Forecasts through uncertainty propagation
result in adequate confidence intervals. However, when missing
values are replaced by the mean of the belief distribution,
further forecasts are overconfident and too many observations
fall outside of 95\% confidence intervals.
